---
title: "M-8 Lab"
author: "Nidhi Bendre"
date: "2023-03-15"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(infer)
library(NHANES)
```


## 2. infer (walk through together)
Calculate a 95% confidence interval for the height of individuals in the NHANES population.

```{r}
# get original sample by filtering out empty values
Sample1 <- NHANES %>%
  filter(!is.na(Height)) %>%
  sample_n(size = 50, replace = FALSE)

#calculate sample mean using dplyr
Sample1 %>% 
  summarise(Xbar = mean(Height))

# calculate sample mean using infer
Xbar <- Sample1 %>%
  specify(response = Height) %>%
  calculate(stat = "mean")
Xbar

# above estimates do not match

#bootstrap confidence interval for μ
boots <- Sample1 %>%
  specify(response = Height) %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "mean")

View(boots)
# shape of distribution is bell curve 
visualize(boots)
```

```{r}
# 95% confidence interval using SE method
ci_se <- get_ci(boots, level = 0.95, type = "se", point_estimate = Xbar)
View(ci_se)
visualize(boots) + shade_confidence_interval(endpoints = ci_se)

# interpretation: We are 95% confident that the true mean height of individuals in the NHANES population is contained in the interval from 149.6452 cm to 164.3788 cm.
```

```{r}
# 95% confidence interval using percentile method
ci_percentile <- get_ci(boots, level = 0.95, type = "percentile")
View(ci_percentile)
visualize(boots) + shade_confidence_interval(endpoints = ci_percentile)

# interpretation: We are 95% confident that the true mean height of individuals in the NHANES population is contained in the interval from 149.4193 cm to 164.0862 cm.
```


## 3. 
Find and interpret a 90% confidence interval for the true population BMI in the NHANES data by taking a sample of size 35.
```{r}
# get original sample by filtering out empty values
sampleBMI <- NHANES %>%
  filter(!is.na(BMI)) %>%
  sample_n(size = 35, replace = FALSE)

# calculate sample mean using infer
Xbar_BMI <- sampleBMI %>%
  specify(response = BMI) %>%
  calculate(stat = "mean")

#bootstrap confidence interval for μ
boots_BMI <- sampleBMI %>%
  specify(response = BMI) %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "mean")

# 95% confidence interval using SE method
ci_se_BMI <- get_ci(boots_BMI, level = 0.90, type = "se", point_estimate = Xbar_BMI)
ci_se_BMI

# interpretation: We are 90% confident that the true mean BMI of individuals in the NHANES population is contained in the interval from 25.38625 to 29.30461.
```


## 4.
Draw a single sample of size 40 from the NHANES dataset and use this to generate a 95% bootstrap confidence interval, based on the percentile method, to estimate the population mean number of hours of sleep per night reported by participants (SleepHrsNight). Interpret this confidence interval.
```{r}
# get original sample by filtering out empty values
sample_sleep <- NHANES %>%
  filter(!is.na(SleepHrsNight)) %>%
  sample_n(size = 40, replace = FALSE)

# calculate sample mean using infer
Xbar_sleep <- sample_sleep %>%
  specify(response = SleepHrsNight) %>%
  calculate(stat = "mean")

#bootstrap confidence interval for μ
boots_sleep <- sample_sleep %>%
  specify(response = SleepHrsNight) %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "mean")

# 95% confidence interval using SE method
ci_percentile_sleep <- get_ci(boots_sleep, level = 0.95, type = "percentile")
ci_percentile_sleep

# interpretation: We are 95% confident that the true mean number of hours of sleep per night of individuals in the NHANES population is contained in the interval from 6.525 hours to 7.45 hours.
```

## 5.
We can also find confidence intervals to give a sense of your uncertainty around point estimates of population proportions from samples.

Use dplyr to create a dataset called drugs that contains the column HardDrugs from the NHANES datasets. Filter out those individuals missing a response for this questions. Sample 25 individuals and store these as Sample3.
```{r}
# dataset created
drugs <- NHANES %>%
  select("HardDrugs")

# missing data filtered and original sample created
Sample3 <- drugs %>%
  filter(!is.na(HardDrugs)) %>%
  sample_n(size = 25, replace = FALSE)
```

Use the infer package to calculate the proportion of individuals in your sample who report using hard drugs in the past year.To calculate the proportion who respond “Yes”, you need to specify the response as HardDrugs and then indicate which answer represents a “success.” Here we want to know what proportion of individuals respond “Yes”, so a "Yes" will be a success. We also need to change the calculate function to calculate the proportion (as opposed to the mean). This is a simple replacement of calculate(stat = "prop").
```{r}
phat <- Sample3 %>%
  specify(response = HardDrugs, success = "Yes") %>%
  calculate(stat = "prop")
phat
# estimate of the true population proportion is 0.28
```

Next, create your 5000 bootstrap resamples, the same way as before, but use the new specify() statement:
```{r}
boots_drugs <- Sample3 %>%
  specify(response = HardDrugs, success = "Yes") %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "prop")
```

Using your estimate, ˆp from above as the “point-estimate”, calculate and interpret the 95% confidence interval using the “standard-error” method.
```{r}
boots_drugs  %>%
  get_ci(level = 0.95, type = "se", point_estimate = phat)

#interpretation: We are 95% confident that the true proportion of NHANES participants reporting yes to hard drugs is contained in the interval from 0.102 to 0.458.
```

Calculate and interpret the 99% confidence interval using the “percentile” method.
```{r}
boots_drugs %>%
get_ci(level = 0.99, type = "percentile")

#interpretation: We are 99% confident that the true proportion of NHANES participants reporting yes to hard drugs is contained in the interval from 0.08 to 0.52.
```

## 6. Difference in Proportions
Suppose you are interested in estimating the difference in the proportion of people who report experimenting with hard drugs among people in their 20s compared to people in their 50s. Let p20 be the true population proportion of those in their 20s who experiment with hard drugs. Let p50 be the true population proportion of those in their 50s who experiment with hard drugs. You are interested in a 90% confidence interval for the difference, p20 − p50.

Draw a sample of 35 people in their 20s and 35 people in their 50s to get an estimate of the difference.
```{r}
Sample6 <- NHANES %>%
  filter(AgeDecade == " 20-29" | AgeDecade == " 50-59") %>%
  filter(!is.na(HardDrugs)) %>%
  group_by(AgeDecade) %>%
  sample_n(35)

Sample6 %>%
  summarise(Xbar = mean(AgeDecade))

boots_age <- Sample6 %>%
  specify(formula = HardDrugs ~ AgeDecade, success = "Yes") %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "diff in props", order = c(" 20-29", " 50-59"))

boots_age %>%
get_ci(level = 0.90, type = "percentile")

#interpretation: We are 90% confident that the true difference of proportions of NHANES participants between ages 20-29 and ages 50-59 reporting yes to hard drugs is contained in the interval from -0.266 to 0.031.
```

## 7. Difference in Means
Oral contraceptive (OC) use has been linked to several blood disorders associated with hormonal changed. Researchers are interested to see if OC increases total cholesterol. Cholesterol levels of 35 women aged 35-50 currently taking OC and 35 women aged 35-50 not currently taking OC were measured. Let μOC be the population average cholesterol level for individuals taking OC. Let μnoOC be the population average cholesterol level for individuals not taking OC.

If OC has no effect on cholesterol, what would you expect μOC − μnoOC to equal?
It should equal 0.

If OC increased cholesterol, would you expect μOC − μnoOC to be positive or negative?
The difference should be positive. 

Suppose the above study results in the following 95% confidence interval for  
μOC − μnoOC : (-1.76, 9.14) mg/dL. Interpret this interval.
We are 95% confidence that the true mean difference in the average cholesterol level reported among 35-50 year-olds who take OC and don't take OC is contained in the interval from -1.76 to 9.14.

Based on the above interval, do you think this OC has an effect on cholesterol?
No, OC does have an effect on cholesterol as 0 is included in the confidence interval. 


